---
title: "BRD model2"
author: "Abdullahi Adinoyi IBRAHIM"
date: "28 August 2018"
output:
  
  html_document: default
  pdf_document: default
---
# Credit Rating Model               
                      
* In this model, we use Colliearity to select statistically significant Featues and then perform Logistic regression on it                

# ================================================        
                           
* Logistic Regression                      
* multicollinearity                                   
                           
# ================================================        

## Some Required packages
```{r}
#install.packages("ROSE")
#Save the required packages in a variable
load_packages <- c("ggplot2","car","Hmisc","ROCR","caret","dummies","caTools","MASS", "gridExtra","gmodels", "ROSE")
#load the packages
lapply(load_packages,require,character.only = TRUE)
```

# Loading the data
```{r}
brd_data <- read.csv("C:\\Users\\abdon\\Documents\\R\\BRD projects guide\\internship\\brd_data1.csv", sep = ",", header =TRUE)
```

# Data prepartion for Modelling
```{r}
#check for NA values
sum(is.na(brd_data))

#check for duplicated values
sum(duplicated(brd_data))

#check the structure and summary
str(brd_data)

#check the summary
#summary(brd_data)
```


# Removing the following columns
```{r}
#brd_data_removed <- brd_data[-c(1,2,3,4)]
brd_data$Count <- NULL
brd_data$GL.229 <- NULL
brd_data$GL.298 <- NULL
brd_data$Code.Reechel. <- NULL
brd_data$Reech.as.On <- NULL
brd_data$Client.Account <- NULL
#===================================
#check for total number of NA
sum(is.na(brd_data))
```


# We Express the continous variables properly 
                       
* They are not properly expressed, we remove the comma
```{r}
#Lets turn all continuous variables to numeric
brd_data$Loan.Amount <- as.numeric(gsub(",","",brd_data$Loan.Amount))
brd_data$GL.A.C <- as.numeric(gsub(",","", brd_data$GL.A.C))
brd_data$Follow.A.C <- as.numeric(gsub(",","", brd_data$Follow.A.C))
brd_data$Settlement.Curr..A.C <- as.numeric(gsub(",","", brd_data$Settlement.Curr..A.C))
#brd_data$PIN.Client <- as.numeric(gsub(",","", brd_data$PIN.Client))
brd_data$Dossier.No <- as.numeric(gsub(",","", brd_data$Dossier.No))
brd_data$Credit.No <- as.numeric(gsub(",","", brd_data$Credit.No))
brd_data$Rate <- as.numeric(gsub(",","", brd_data$Rate))
brd_data$Del.Int..Rate <- as.numeric(gsub(",","", brd_data$Del.Int..Rate))
#brd_data$Loan.Duration <- as.numeric(gsub(",","", brd_data$Loan.Amount))
brd_data$Disbursed <- as.numeric(gsub(",","", brd_data$Disbursed))
brd_data$To.Be.Disbursed <- as.numeric(gsub(",","", brd_data$To.Be.Disbursed))
brd_data$Reimbursed.Amount <- as.numeric(gsub(",","", brd_data$Reimbursed.Amount))
brd_data$Debit.Balance <- as.numeric(gsub(",","", brd_data$Debit.Balance))
brd_data$Solde.Comptable <- as.numeric(gsub(",","", brd_data$Solde.Comptable))
brd_data$Int...com.issued <- as.numeric(gsub(",","", brd_data$Int...com.issued))
brd_data$Int...com.paid <- as.numeric(gsub(",","", brd_data$Int...com.paid))
brd_data$Int...Com.unpaid <- as.numeric(gsub(",","", brd_data$Int...Com.unpaid))
brd_data$Int.balance...com <- as.numeric(gsub(",","", brd_data$Int.balance...com))
brd_data$Total.Risk <- as.numeric(gsub(",","", brd_data$Total.Risk))
brd_data$Principal.unpaid <- as.numeric(gsub(",","", brd_data$Principal.unpaid))
brd_data$Norm..Interest <- as.numeric(gsub(",","", brd_data$Norm..Interest))
brd_data$Delay..Interest <- as.numeric(gsub(",","", brd_data$Delay..Interest))
brd_data$Comms <- as.numeric(gsub(",","", brd_data$Comms))
brd_data$Activity <- as.numeric(gsub(",","", brd_data$Activity))
brd_data$X.PRI.INT..Next.Due.Instalment <- as.numeric(gsub(",","", brd_data$X.PRI.INT..Next.Due.Instalment))
brd_data$Principal <- as.numeric(gsub(",","", brd_data$Principal))
brd_data$Interest <- as.numeric(gsub(",","", brd_data$Interest))
brd_data$Grand.total <- as.numeric(gsub(",","", brd_data$Grand.total))
brd_data$suspended.int <- as.numeric(gsub(",","", brd_data$suspended.int))
brd_data$Collateral <- as.numeric(gsub(",","", brd_data$Collateral))
brd_data$Provisions <- as.numeric(gsub(",","", brd_data$Provisions))
```

```{r}
brd_data <- na.omit(brd_data) #incase of NA's introduced by coercion
```
                        
* Since there are no more NA's, Now extract some features and check for outliers

# Selecting Features
```{r}
# Extract the following variables variables only
brd_data <- brd_data[c("Loan.Amount",
                                 "GL.A.C",
                                 "Follow.A.C",
                                 "Settlement.Curr..A.C",
                                 "Dossier.No",
                                 "Credit.No",
                                 "Rate",
                                 "Del.Int..Rate",
                                 "Loan.Duration",
                                 "Disbursed",
                                "To.Be.Disbursed",
                                "Reimbursed.Amount",
                                "Debit.Balance",
                                "Solde.Comptable",
                                "Int...com.issued",
                                "Int...com.paid",
                                "Int...Com.unpaid",
                                "Int.balance...com",
                                "Total.Risk",
                                "Principal.unpaid",
                                "Norm..Interest",
                                "Delay..Interest",
                                "Comms",
                                "Activity",
                                "X.PRI.INT..Next.Due.Instalment",
                                "Principal",
                                "Interest",
                                "Grand.total",
                                "suspended.int",
                                "Collateral",
                                "Provisions",
                                "Class.of.Risk",
                                "Monitored.By",
                                "District",
                                "BRD.Sector",
                                "Sector",
                                "DIVISION"
)]
```

# ================================================        
# creating a response variable
# ================================================        

```{r}
brd_data$status <- ifelse(brd_data$Class.of.Risk >= 3, 1,0) 
brd_data$status <- factor( brd_data$status, levels = c(0,1))
#default=1
#non default = 0
```

# ================================================        
# Univariate analysis for the **categorical** variables                
# ================================================        


## **Class.of.Risk** column                         
```{r}
ggplot(brd_data) +
  geom_bar(aes(x=Class.of.Risk,fill = "gray")) + 
  theme(axis.text.y = element_text(size = rel(.8)) )
```

## **Monitored.By** column
```{r}
attach(brd_data)
ggplot(brd_data) +
  geom_bar(aes(x=Monitored.By), fill = 'gray') + 
  coord_flip()+
  theme(axis.text.y = element_text(size = rel(.8)) )
```

## **BRD.Sector** column
```{r}
ggplot(brd_data) +
  geom_bar(aes(x=BRD.Sector),fill = 'gray') + 
  coord_flip()+
  theme(axis.text.y = element_text(size = rel(.8)) )
```

## **sector** column
```{r}
ggplot(brd_data) +
  geom_bar(aes(x=Sector),fill = 'gray') + 
  coord_flip()+
  theme(axis.text.y = element_text(size = rel(.8)) )
```

## **DIVISION** column
```{r}
ggplot(brd_data) +
  geom_bar(aes(x=DIVISION),fill = 'gray') + 
  theme(axis.text.y = element_text(size = rel(.8)) )
```

# Bivariate plot
## **District** & **Loan.Amount** columns 
```{r}
ggplot(brd_data) +
  geom_bar(aes(x=District,y=Loan.Amount),stat = "identity", fill = 'gray') + 
  coord_flip()+
  theme(axis.text.y = element_text(size = rel(.8)) )
```

# ================================================        
# selection of Continous Variables/Features
# ================================================        

```{r}
#brd_data <- brd_data[c(-23,-33,-34,-35,-36,-37)]
brd_data <- brd_data[c(-23,-33,-34,-35,-37)]
```

# No more NA's in the Data, Now we check for outliers
## outlier treatment for all **Continous** variables

```{r}
#column = Settlement.Curr..A.C
#check for outliers in numeric variables, if any found than we do capping and flooring
options(scipen = 999) #Express exponentials properly
boxplot.stats(brd_data$Settlement.Curr..A.C) # No Outlier
#boxplot(na.omit(brd_data1$Settlement.Curr..A.C), main = "boxplot for 'Settlement.Curr..A.C' ")
plot(brd_data$Settlement.Curr..A.C)
#boxplot.stats(brd_data$Settlement.Curr..A.C)$out 
```


## column = Settlement.Curr..A.C
```{r}
#column = Settlement.Curr..A.C
#check for outliers in numeric variables, if any found than we do capping and flooring
options(scipen = 999) #transfrom from exponent form
boxplot.stats(brd_data$Follow.A.C) #no outlier
boxplot(brd_data$Follow.A.C, main = "boxplot for 'follow_A.C' ")
#boxplot.stats(brd_data1$Follow.A.C)$out
```

## column = Loan.Amount
```{r}
# column = Loan.Amount
#check for outliers in numeric variables, if any found than we do capping and flooring
options(scipen = 999)

boxplot.stats(brd_data$Loan.Amount) #103 outliers detected
boxplot(brd_data$Loan.Amount, main = "boxplot for 'Loan.Amount' ")
plot(brd_data$Loan.Amount) #many visible outliers
#================================
#outlier
#a <- quantile(brd_data$Loan.Amount)       #IRQ = a['75%'] - a['25%'] 
outliers_upperlimit <- quantile(brd_data$Loan.Amount, 0.75) + 1.5 * IQR(brd_data$Loan.Amount)
index.outliers.loan <- which(brd_data$Loan.Amount >= outliers_upperlimit | brd_data$Loan.Amount >= 102136587 ) 
brd_data <- brd_data[-index.outliers.loan, ] #Removing observations

plot(brd_data$Loan.Amount)#check the outliers
boxplot(brd_data$Loan.Amount,main = "boxplot without outlier ") # no outliers
boxplot.stats(brd_data$Loan.Amount) # no more outliers
```

## column = Collateral
```{r}
#column = Collateral
#check for outliers in numeric variables, if any found than we do capping and flooring
options(scipen = 999)

boxplot.stats(brd_data$Collateral)
boxplot(brd_data$Collateral, main = "boxplot for 'Collateral' ") #many outliers
plot(brd_data$Collateral)
#min(boxplot.stats(brd_data$Collateral)$out) # No Outlier

# outlier treatmernt
outliers_upperlimit1 <- quantile(brd_data$Collateral, 0.75) + 1.5 * IQR(brd_data$Collateral) # upper_limit = 
index.outliers.collateral <- which(brd_data$Collateral >= outliers_upperlimit | brd_data$Collateral >= 125961 ) #470 outliers
brd_data <- brd_data[-index.outliers.collateral, ] #Removing observations

plot(brd_data$Collateral, main = "plot without outliers")#check the outliers
boxplot(brd_data$Collateral, main = "boxplot without outlier") # no outliers
boxplot.stats(brd_data$Loan.Amount) # no more outliers

```

## column = grant.total
```{r}
#column = grant.total
#check for outliers in numeric variables, if any found than we do capping and flooring

boxplot.stats(brd_data$Grand.total)
boxplot(brd_data$Grand.total, main = "boxplot for 'Grand.total' ")
plot(brd_data$Grand.total, main = "plot with outliers")
#min(boxplot.stats(brd_data$Grand.total)$out) # No Outlier

# outlier treatmernt
outliers_upperlimit2 <- quantile(brd_data$Grand.total, 0.75) + 1.5 * IQR(brd_data$Grand.total) # upper_limit = 
index.outliers.Grand_total <- which(brd_data$Grand.total >= outliers_upperlimit | brd_data$Grand.total >= 76870 ) #5 outliers
brd_data <- brd_data[-index.outliers.Grand_total, ] #Removing observations

plot(brd_data$Grand.total, main = "plot without outliers")#double check the outliers
boxplot(brd_data$Grand.total, main = "boxplot without outlier", xlab = "Grand.total") # no outliers
boxplot.stats(brd_data$Grand.total) # no more outliers
```

## column = principal
```{r}
#column = principal
#check for outliers in numeric variables, if any found than we do capping and flooring
boxplot.stats(brd_data$Principal) #49 outliers
boxplot(brd_data$Principal, main = "boxplot for 'Principal' ")
plot(brd_data$Principal)
# outlier treatmernt
outliers_upperlimit3 <- quantile(brd_data$Principal, 0.75) + 1.5 * IQR(brd_data$Principal) # upper_limit = 
index.outliers.principal <- which(brd_data$Principal >= outliers_upperlimit3 | brd_data$Principal >=  337283 ) #50 outliers
brd_data <- brd_data[-index.outliers.principal, ] #Removing observations

plot(brd_data$Principal, main = "plot without outliers")#double check the outliers
boxplot(brd_data$Principal, main = "boxplot without outlier",xlab = "Principal") # no outliers
#boxplot.stats(brd_data$principal) # no more outliers
```

## column = Total.risk
```{r}
#column = Total.Risk
#check for outliers in numeric variables, if any found than we do capping and flooring
boxplot.stats(brd_data$Total.Risk) #No outliers
boxplot(brd_data$Total.Risk, main = "boxplot for 'Total.Risk' ")
plot(brd_data$Total.Risk)
```

# ---------------------------
# MULTICOLLINEARITY
# -----------------------------
```{r}
# to display the correlations
#cor(brd_data[-32])

```
           
* We check for correlation and remove features with high correlation
                        
# we shall cutoff variables with high correlation
```{r}
#highly_Cor_brd <- colnames(brd_data)[findCorrelation(abs(cor(brd_data[-32])), cutoff = 0.75, verbose = TRUE)]
#print(highlyCor)

```

# **Removing the features**
```{r}
#**Removing the columns**
#brd_data_okay <- brd_data[,which(!colnames(brd_data) %in% highly_Cor_brd[-32])]
#----------------------------------------------------
# Now we have low correlation
#cor(brd_data_okay)
```

# =====================================================        
# splitting the data into **train** and **test** sample
# =====================================================        
                     
* this step is necessary, we need to split the data into two parts, train(70%) and test(30). then we build our model from the train sample and use the test_sample to evaluate the model performance
```{r}
# creating a response variable as a factor of 0 and 1
brd_data$status <- factor( brd_data$status, levels = c(0,1))
table(brd_data$status) #check data type
#default=1
#non default = 0

# =================
set.seed(100)
sample <- sample.split(brd_data$status, 0.7)
train <- subset(brd_data, sample == TRUE)
test <- subset(brd_data, sample == FALSE)
```


```{r}
prop.table(table( brd_data$status ))
# non-default =89.3% and default = 10.7% 
#dim(train)
```


# balance the data
```{r}
#table(train$status)
# the response variable is bias
# we have to baance the data
data.rose <- ROSE(status ~ ., data = train, seed = 1)$data
prop.table( table(data.rose$status))
# default=50.5  and non-default = 49.5%
```
                          
* Now we have balance the data. we can now develop the model                 


                                                     
# ================================================        
# Logistic Regression Model              
# ================================================        
```{r}
#names(data.rose)
#m1 <- glm(formula = status ~ ., data =  data.rose[-1], family = binomial)
Logitmodel <- glm(formula = status ~ Loan.Amount + Activity + Rate + Collateral   , data =  data.rose, family = binomial)
summary(Logitmodel)
```


# Prediction
```{r}
test$pred <- predict(Logitmodel, newdata = test, type = "response")
hist(test$pred, main = "histogram for predcited")
```

## Evaluate the model
```{r}
#test$pred <- predict(m1, newdata = test, type = "response")
predic <- prediction(test$pred,test$status)
perf <- performance(predic, "tpr", "fpr")
plot(perf)

#accuracy.meas(test$status, test$pred)
```

# the prediction
```{r}
test$cut_off <- ifelse(test$pred >= .6, 1, 0 )
table(test$status, test$cut_off)

# numbers and their percentage
#CrossTable(test$status, test$cut_off,prop.r = TRUE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE)
```

## SUmmary of the prediction                             
```{r}
confusionMatrix(as.factor( test$status),as.factor( test$cut_off))
```
                  
* this function gives the model accuracy, the specificity and specitivity

## overall accuracy of the model in percentage
```{r}
sum(diag(table(test$status, test$cut_off )))/ nrow(test) * 100 
```


```{r}
colAUC(test$pred,test$status, plotROC = TRUE)
```

# Rating scale
* Now we are going to rate the borrowers according to the rating probability criteria
```{r}
test$Rating <-           ifelse(test$pred >=0.0 & test$pred < .01, "AAA",  
                         ifelse(test$pred >= .01 & test$pred < .02, "AA",
                         ifelse( test$pred >= 0.02 & test$pred < 0.05, "A",
                         ifelse( test$pred >= 0.05 & test$pred < .10, "BBB",
                         ifelse( test$pred >= 0.1 & test$pred < .20, "BB",
                         ifelse( test$pred >= .20 & test$pred < .40, "B",
                         ifelse( test$pred >= .40 & test$pred < 1, "C",
                         ifelse( test$pred >= 1 & test$pred <= 1, "D", test$pred ))))))))
#=======================
table(test$Rating)
```
                                                 
# Lets round the **pred** to 2 decimal places                
             
## Since our predicted-values are more than 2 decimals and our rating criteria is in 2 decimal places
```{r}
test$aprrox <- round(test$pred,digits=2)
# ---------------------------------
test$Rating2 <-          ifelse(test$aprrox >=0.0 & test$aprrox < .01, "AAA",  
                         ifelse(test$aprrox >= .01 & test$aprrox < .02, "AA",
                         ifelse( test$aprrox >= 0.02 & test$aprrox < 0.05, "A",
                         ifelse( test$aprrox >= 0.05 & test$aprrox < .10, "BBB",
                         ifelse( test$aprrox >= 0.1 & test$aprrox < .20, "BB",
                         ifelse( test$aprrox >= .20 & test$aprrox < .40, "B",
                         ifelse( test$aprrox >= .40 & test$aprrox < 1, "C",
                         ifelse( test$aprrox >= 1, "D", test$aprrox ))))))))
#=======================
table(test$Rating2)
CrossTable(test$Rating2)
```

# Visualize final rating
```{r}
ggplot(test) + 
  geom_bar(aes(x=Rating2),fill = 'gray') + 
  coord_flip() +
  theme(axis.text.y = element_text(size = rel(.8))) +ggtitle("Rating classes and its frequecy")
```

# ================================================        
# Conclusion
# ================================================        
         
* (1) from above, we have been able to determine the following (1) probability of default and (2 )rate the borrowers
              
* (2) to make a prediction, the following feature must be present in our dataset: ( Loan.Amount,  Activity,  Rate,  Collateral )  
        
* (3) I balanced the data because Had I just blindly predicted all the data points as non-default, I would achieve an accuracy              percentage of 90%. Which sounds pretty high. But obviously that is flawed. What matters is how well you predict the default           classes.So that requires the default and non-default classes are balanced AND on top of that I need more refined accuracy             measures and model evaluation metrics to improve my prediction model.
                                   
     
# =================================================        
# =================================================                                           
# Example 1 

```{r}
library(tidyverse)
data3 <- read.csv("C:\\Users\\abdon\\Documents\\R\\BRD projects guide\\internship\\Submit\\data6.csv", header = TRUE, sep = ",")
#data3 <- data3[c("Loan.Amount", "Activity", "Rate", "Collateral", "Sector", ")]

data3$Loan.Amount <- as.numeric(gsub(",","",data3$Loan.Amount))
data3$Activity <- as.numeric(gsub(",","",data3$Activity))
data3$Rate <- as.numeric(gsub(",","",data3$Rate))
data3$Collateral <- as.numeric(gsub(",","",data3$Collateral))
#data3$Principal <- as.numeric(gsub(",", "", data3$Principal))
# --------------------------
data3$pred <- Logitmodel %>% predict(newdata = data3, type = "response")
data3$status <- as.factor(ifelse(data3$pred >= .6, 1,0))
# Rating
data3$aprrox <- round(data3$pred,digits=2)
data3$Ratings <-        ifelse( data3$aprrox >=0.0 & data3$aprrox < .01, "AAA",  
                         ifelse( data3$aprrox >= .01 & data3$aprrox < .02, "AA",
                         ifelse( data3$aprrox >= 0.02 & data3$aprrox < 0.05, "A",
                         ifelse( data3$aprrox >= 0.05 & data3$aprrox < .10, "BBB",
                         ifelse( data3$aprrox >= 0.1 & data3$aprrox < .20, "BB",
                         ifelse( data3$aprrox >= .20 & data3$aprrox < .40, "B",
                         ifelse( data3$aprrox >= .40 & data3$aprrox < 1, "C",
                         ifelse( data3$aprrox >= 1, "D", data3$aprrox ))))))))
# --------------------------
table(data3$Ratings)
# searching for a row
#data3[data3$Loan.Amount == 20968092 ,]
```


```{r}
# to knit pdf
# install.packages("rmarkdown")
# devtools::install_github('yihui/tinytex')
data3$original <- as.factor( ifelse(brd_data [brd_data$Rate == data3$Rate && brd_data$Class.of.Risk >= 3, 1,0]))
```




